The Topic Browser\\An Interactive Tool for Browsing Topic Models

### 2010-10-21 masters_thesis/browser_nips/paper.tex
Finished a draft of the browser paper with pictures

When browsing through topics, the user can filter the topic list by coherence
to eliminate from the view topics that are mostly meaningless, and sort by
document entropy to find topics that were used widely (or narrowly, if that is
desired) throughout the corpus.  She could also filter by document or
attribute, seeing only topics that were used in a particular document or by a
particular author.  Documents can similarly be filtered and sorted, showing
only documents that contain tokens with a particular topic, or with a
particular value of an attribute.

@@

The topics in a topic model cannot be fully interpreted when completely
separated from their context.  Thus we display the top ten words in the topic
inside of their context.  We select a random token of each word type labeled
with the topic and show up to 50 characters on either side of the token,
keeping words intact.  We also allow the user to cycle through contexts to gain
a broader view of how each word is actually used in that topic in the corpus.

@@

The browser provides two main ways of getting more information about topics in
the model.  The first is showing documents and attribute values which have
either the highest number of tokens labeled with that topic or the highest
proportion (by percentage) of the given topic.  This allows the user to quickly
find documents that best demonstrate what the topic captured.  Showing the top
values for a given attribute (such as authors for the attribute ``Author'')
gives the user an idea of how focused the topic is and often gives interesting
information about the corpus being browsed (for example, in Figure
\ref{fig:something}, we see that Democrats spoke more often about education
than Republicans did, at least in the topic shown).

The other way that we give more information about topics is by showing similar
topics.  We find similar topics either by looking at the documents that the
topics show up in or by the words that the topics use.  Looking at similar
topics by document shows topics that are commonly used together in the corpus,
and looking by word shows topics that have similar words, possibly because the
two topics really should have been one topic.

@@

When simply browsing through the documents, we provide sorting and filtering
methods on the list of documents, as mentioned previously.  When looking at a
particular document, we show basic information about the document, its text,
the topic distribution in the document, and similar documents based on that
distribution.  When looking at the document in the context of a topic, we also
highlight the tokens in the document that were labeled with that topic.  All of
this has been done previously, aside from the sorting and filtering of the
documents, and in fact constitutes the entirety of most previous corpus
browsers based on topic models---this functionality is all that is generally
provided.

We also provide views of individual words in the corpus.  When viewing a word
in the context of a topic, the user can see all uses of the word in that topic
in the corpus, with context taken from their corresponding documents.  The user
can also view words independently with a search-like interface, seeing topics
and documents in which the word appears most frequently.  While providing basic
functionality, however, the search interface leaves much to be desired, as only
single words can currently be searched for.  We plan on expanding that to
phrases.

The user can also look at aggregated information for values of an attribute
(i.e., a particular author or year in the documents), combining all of the
topic and word counts for all documents with the given attribute.  This view is
also at present somewhat limited, showing only what topics and words are used
most frequently by the collection of documents with that attribute.

@@

We currently include two kinds of plots in our topic browser, with plans to
implement many more.  The first is showing trends for topics over the values of
an attribute (such as date, or author), useful for corpus browsing.  This kind
of plot has been used in visualizing topic models almost since their
introduction~\cite{griffiths-2004-finding-scientific-topics}.  Our topic
browser allows the user to interactively generate these trend plots over any
attribute for any topic or combination of topics of the documents in the
corpus.

The second kind of plot that we include is more useful for analyzing and
understanding the behavior of the topic model itself.  We allow the user to
plot two topic metrics against each other and compute a linear regression.
This allows the user to see some interesting properties of the topic metrics,
such as the fact that document entropy seems to correlate with the logarithm of
the number of tokens in the topic, and that coherence does not seem to
correlate with any other topic metric.  The user can also find outliers, such
as topics with low document entropy but a high token count, that can then be
examined in the topic page. (TODO: use Obama vs. McCain as example here)

>>>>

When browsing through topics, the user can filter the topic list by coherence
to eliminate from the view topics that are mostly meaningless, and sort by
document entropy to find topics that were used widely (or narrowly, if that is
desired) throughout the corpus.  For example, Figure~\ref{fig:sidebar} shows
that the family was one of the most consistent themes in the speeches.

\begin{figure}
  \centering
  \includegraphics[width=.34\textwidth]{sidebar}
  \caption{The navigational sidebar showing topics sorted by dispersion
  throughout the corpus and filtered by coherence}
  \label{fig:sidebar}
\end{figure}

The user can also filter by document or attribute, seeing only topics that were
used in a particular document or by a particular author.  Documents can
similarly be filtered and sorted, showing only documents that contain tokens
with a particular topic, or with a particular value of an attribute.

@@

The topics in a topic model cannot be fully interpreted when completely
separated from their context.  Thus we display the top ten words in the topic
inside of their context.  We select a random token of each word type labeled
with the topic and show up to 50 characters on either side of the token,
keeping words intact.  We also allow the user to cycle through contexts to gain
a broader view of how each word is actually used in that topic in the corpus.
When viewing the topic about troops in Iraq, for example (see
Figure~\ref{fig:context}), one can see that most often when the word ``troops''
is used in this topic, it is in reference to bringing the troops home, and when
``end'' is used, it is not used in the context of ``end the war'' as often as
one might expect.

\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{words_in_context}
  \caption{The top ten words in the topic shown with a random context selected
  from the corpus.}
  \label{fig:context}
\end{figure}

@@

The browser provides two main ways of getting more information about topics in
the model.  The first is showing documents and attribute values which have
either the highest number of tokens labeled with that topic or the highest
proportion (by percentage) of the given topic.  This allows the user to quickly
find documents that best demonstrate what the topic captured.  Showing the top
values for a given attribute (such as authors for the attribute ``Author'')
gives the user an idea of how focused the topic is and often gives interesting
information about the corpus being browsed.  For example, in Figure
\ref{fig:top-values}, we see that Democrats spoke more often about education
than Republicans did, at least in the topic shown.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{top_values}
  \caption{Part of the overall topic view, looking at a topic about education.
  Note particularly the top documents for the topic and the top values for the
  attribute ``party,'' shown at the bottom.}
  \label{fig:top-values}
\end{figure}

The other way that we give more information about topics is by showing similar
topics.  We find similar topics either by looking at the documents that the
topics show up in or by the words that the topics use.  Looking at similar
topics by document shows topics that are commonly used together in the corpus,
and looking by word shows topics that have similar words, possibly because the
two topics really should have been one topic.  When looking at a topic about
health care, the user can see that other topics used together with the health
care topic include topics about the cost of medication and the quality of care
(see Figure~\ref{fig:similar}).

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{similar_topics}
  \caption{Another look at the overall topic view, featuring different parts of
  the view.  Note particularly the similar topics box near the top right.}
  \label{fig:similar}
\end{figure}

@@

When simply browsing through the documents, we provide sorting and filtering
methods on the list of documents, as mentioned previously.  When looking at a
particular document, we show basic information about the document, its text,
the topic distribution in the document, and similar documents based on that
distribution.  When looking at the document in the context of a topic, we also
highlight the tokens in the document that were labeled with that topic.  For
example, the user might be curious to see a document that best demonstrates the
health care topic mentioned above.  Clicking on the top document in the document list brings the user to the view shown in Figure~\ref{fig:document}.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{document}
  \caption{The document view, looking at a speech by Hillary Clinton in the
  context of a topic about health care (the rest of the document is cut off in
  this screenshot).}
  \label{fig:document}
\end{figure}

These document visualizations been done previously, and in fact the view in
Figure~\ref{fig:document} constitutes the entirety of most previous corpus
browsers based on topic models.  This functionality is all that is generally
provided in other browsers, though it is just a small part of ours.

We also provide views of individual words in the corpus.  When viewing a word
in the context of a topic, the user can see all uses of the word in that topic
in the corpus, with context taken from their corresponding documents.  The user
can also view words independently with a search-like interface, seeing topics
and documents in which the word appears most frequently.  The user examining
the health care topic might be curious where else the word ``health'' was used
in the topics and the corpus.  Figure~\ref{fig:word} shows the result of using
our word search to answer that question.  While providing basic functionality,
however, the search interface leaves much to be desired, as only single words
can currently be searched for.  We plan on expanding that to phrases.

\begin{figure}
  \centering
  \includegraphics[width=.72\textwidth]{word}
  \caption{The word search functionality, showing the results for searching for
  the word ``health.''}
  \label{fig:word}
\end{figure}

The user can also look at aggregated information for values of an attribute
(i.e., a particular author or year in the documents), combining all of the
topic and word counts for all documents with the given attribute.  This view is
also at present somewhat limited, showing only what topics and words are used
most frequently by the collection of documents with that attribute.
Figure~\ref{fig:attribute} shows us, among other things, that one of Barack
Obama's top topics was change in politics.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{attribute}
  \caption{The attribute view, showing aggregated information for all speeches
  given by Barack Obama.}
  \label{fig:attribute}
\end{figure}

@@

We currently include two kinds of plots in our topic browser, with plans to
implement many more.  The first is showing trends for topics over the values of
an attribute (such as date, or author), useful for corpus browsing.  This kind
of plot has been used in visualizing topic models almost since their
introduction~\cite{griffiths-2004-finding-scientific-topics}.  Our topic
browser allows the user to interactively generate these trend plots over any
attribute for any topic or combination of topics of the documents in the
corpus.  Our user exploring health care topics may want to view how much each
candidate spoke about health care.  We saw already that there were three
related topics that mentioned health care, so the user views all of them
together in a histogram, shown in Figure~\ref{fig:trend-plot}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{trend_plot}
  \caption{A plot of topics over attributes, showing the use of three health
  care-related topics across candidates.}
  \label{fig:trend-plot}
\end{figure}

The second kind of plot that we include is more useful for analyzing and
understanding the behavior of the topic model itself.  We allow the user to
plot two topic metrics against each other and compute a linear regression.
This allows the user to see some interesting properties of the topic metrics,
such as the fact that document entropy seems to correlate with the logarithm of
the number of tokens in the topic, and that coherence does not seem to
correlate with any other topic metric.  The user can also find outliers, such
as topics with low document entropy but a high token count, that can then be
examined in the topic page.

An interesting application of these topic metric plots occurs when the metrics
include how consistently each candidate spoke about each topic.
Figure~\ref{fig:metric-plot} shows a plot of topics, comparing John McCain's
use of the topic to Barack Obama's use of the topic.  Topics in the upper left
were topics unique to Obama, and topics in the lower right were unique to
McCain.  Topics in the middle (near the value of 2 for each candidate) were
somewhat shared between the two.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{metric_plot}
  \caption{A topic metric comparison plot.  The metrics used were how
  consistently Barack Obama and John McCain used each topic. The plot shows
  topics that were unique to Obama or McCain and topics they shared.}
  \label{fig:metric-plot}
\end{figure}
